### 2020-06-24 neutronics.tex
Typo in nts chapter, complete dnp section in ss chapter, fix figures

The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
temperature of the reactor. We can raise or lower the average core temperature
at steady state to meet the design specifications for the inlet and outlet
temperatures by adjusting the fissile inventory and $k_{\text{eff}}$. In
transient analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are
important as they dictate the duration and magnitude of the reactor
response to a transient initiator. We calculated $\beta_{\text{eff}}$ by
taking the relative difference between the $k_{\text{eff}}$ values with and
without \glspl{DNP} in Table \ref{table:keff}. The $\beta_{\text{eff}}$ values
at 973 K, shown in Table \ref{table:betaeff}, are in excellent agreement with
a 4.43 pcm discrepancy.

\begin{table}[htb!]
	\centering
	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
	\begin{tabular}{l S S}
		\toprule
		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Difference wrt Serpent [pcm]}
		\\
		\midrule
		{Serpent} & 304.08 \pm .81 & {-}\\
		{Moltres} & 299.65 \pm .20 & 4.43\\
		\bottomrule
	\end{tabular}
	\label{table:betaeff}
\end{table}

>>>>

The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
temperature of the reactor. We can raise or lower the average core temperature
at steady state to meet the design specifications for the inlet and outlet
temperatures by adjusting the fissile inventory and $k_{\text{eff}}$. In
transient analysis, the delayed neutron fraction $\beta$ and reactivity
coefficients $\alpha$ are
important as they dictate the duration, shape, and magnitude of the reactor
response to a transient initiator. The difference between $\beta$ and
$\beta_{\text{eff}}$ is that $\beta$ is the physical delayed neutron fraction
while $\beta_{\text{eff}}$ is the delayed neutron fraction weighted on the
adjoint neutron flux. We will compare the $\beta$ value from Moltres to the
$\beta_{\text{eff}}$ value from Serpent because Moltres currently lacks a
adjoint calculation capability. We calculated $\beta$ by
taking the relative difference between the $k_{\text{eff}}$ values with and
without \glspl{DNP} in Table \ref{table:keff}. The $\beta$ and
$\beta_{\text{eff}}$ values at 973 K, shown in Table \ref{table:betaeff}, are
in good agreement with a 4.43 pcm discrepancy.

\begin{table}[htb!]
	\centering
	\caption{$\beta_{\text{eff}}$ and $\beta$ values from Serpent and Moltres,
	respectively, at 973 K.}
	\begin{tabular}{l S S}
		\toprule
		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Difference wrt Serpent [pcm]}
		\\
		\midrule
		{Serpent} & 304.08 \pm .81 & {-}\\
		{Moltres} & 299.65 \pm .20 & 4.43\\
		\bottomrule
	\end{tabular}
	\label{table:betaeff}
\end{table}

### 2020-06-24 steadystate.tex
Typo in nts chapter, complete dnp section in ss chapter, fix figures

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi/TUDelft
models. The results from Moltres show good qualitative agreement with the
Polimi and TUDelft models \cite{aufiero_development_2014}; we observe similar
flow and hotspot features in all three models. There is a large recirculation
region near the blanket tank walls arising from turbulent flow. Inertial
forces dominate over viscous forces to form this large eddy. The figure also
shows two relatively stagnant regions along the central axis at the top and
bottom of the core. Temperature hotspots form in these regions of
recirculation as convection is
the dominant heat transfer mechanism. The maximum temperature from Moltres,
1275 K near the bottom of the recirculation zone, is closer to the maximum
temperature in the Polimi model ($\approx$ 1300 K) than the TUDelft model
($\approx$ 1200K). The minimum temperature is 922 K at the inlet. Figure
\ref{fig:flow} provides an alternate view of the flow profile through flow
streamlines superimposed on the velocity magnitude distribution. The fastest
salt flow occurs at the inlet, outlet, and at core half-height approximately
0.40 m away from the central axis.

@@

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g1}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g2}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g3}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g4}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g5}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g6}
    \end{subfigure}
    \caption{Neutron flux distributions in the core for neutron energy groups
    1 to 6. The y and x axes represent height and radius values of the core
    relative to the entire reactor geometry.}
    \label{fig:neutronflux}
\end{figure}

@@

\begin{table}[t!]
	\centering
	\caption{Peak neutron flux values from Moltres (this paper), COMSOL
	\cite{fiorina_molten_2013}, and OpenFOAM \cite{aufiero_development_2014}
	models. The COMSOL and OpenFOAM values were calculated with uniform
	temperature distributions.}
	\begin{tabular}{l S}
		\toprule
		{Model} & {Peak Neutron Flux [$\times 10^{15}$ cm$^{-2}$ s$^{-1}$]}
		\\
		\midrule
		{Moltres (This paper)} & 9.80\\
		{COMSOL} & 8.6 \\
		{OpenFOAM} & 9.0 \\
		\bottomrule
	\end{tabular}
	\label{table:peak-flux}
\end{table}

As mentioned before, the \glspl{DNP} are mobile in \glspl{MSR} and their
distributions do not directly correspond to the neutron flux distributions.
The location where the \glspl{DNP} decay and emit neutrons impacts their
neutron
importance depending on their proximity to fissile and parasitic isotopes.
Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
groups. The
precursors from the shortest-lived group (Group 8) predominantly decay within
the core as their half-lives are shorter than the time it takes to reach the
outlet while the precursors from the longest-lived group (Group 1) are
effectively uniformly distributed due to their long half-lives relative to the
average circulation time. For the longer-lived groups, the mesh elements with
relatively high and low \gls{DNP} concentrations near the outlet and inlet,
respectively, suggest that the mesh is too coarse. Thus, we would recommend
further refinement for future work involving similar geometries.

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre1}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre2}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre3}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre4}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre5}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre6}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre7}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre8}
    \end{subfigure}
    \caption{\gls{DNP} distributions in the core for \gls{DNP} groups
    1 to 8 (from left to right, top to bottom). Refer to Figure \ref{fig:neutronflux} for the height and radius
    scales on the y and x axes, respectively. Note the different scales for
    each distribution.}
    \label{fig:dnp}
\end{figure}

\subsection{In-Core \texorpdfstring{$\beta_{\text{eff}}$}{beta_eff}}

An important safety parameter for \glspl{MSR} is the actual in-core
$\beta_{\text{eff}}$ value. This value represents the actual effective delayed
neutron fraction in \glspl{MSR} after accounting for the loss of delayed
neutrons due to out-of-core \gls{DNP} decay. Table * shows

>>>>

Figure \ref{fig:flow-temp} shows the temperature and velocity fields of the
fuel salt in the core at steady state from Moltres and the Polimi and TUDelft
models. Figure \ref{fig:flow} provides an alternate view of the flow profile
through flow streamlines superimposed on the velocity magnitude distribution.
The results from Moltres show good qualitative agreement with the
Polimi and TUDelft models \cite{fiorina_modelling_2014}; we observe similar
flow and hotspot features in all three models. Furthermore, the highest salt
velocities in all three models occur at the inlet, outlet, and at core
half-height approximately 0.40 m away from the central axis. There is a large
recirculation region near the blanket tank walls arising from turbulent flow.
Inertial forces dominate over viscous forces to form this large eddy. The main
difference between Moltres and Fiorina et al.'s models is the flow profile
near the central axis at the top and bottom of the core.
The Polimi and TUDelft models predict relatively stagnant flow in these
regions without recirculation. Moltres, on the other hand, predicts explicit
recirculating flow in these regions. This is due to our constant turbulent
viscosity approximation in Moltres. The k-$\epsilon$ turbulence models in the
Polimi and TUDelft models predict that the turbulent viscosity in these
regions is as high as 100 Pa s, much higher than our 40 Pa s approximation.

Nevertheless, similar temperature hotspots form in these regions of
recirculation and stagnation as convection is the dominant heat transfer
mechanism. The maximum temperature from Moltres, 1275 K near the bottom of the
large recirculation zone, is closer to the maximum temperature in the Polimi
model ($\approx$ 1300 K) than the TUDelft model ($\approx$ 1200K). Similarly,
we observe cooler temperatures in high-velocity regions. The minimum
temperature is 924 K at the inlet.

@@

\begin{figure}[b!]
    \centering
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g1}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g2}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g3}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g4}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g5}
    \end{subfigure}
    \begin{subfigure}[t]{.325\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-g6}
    \end{subfigure}
    \caption{Neutron flux distributions in the core for neutron energy groups
    1 to 6. The y and x axes represent height and radius (in cm) of the core
    relative to the entire reactor geometry.}
    \label{fig:neutronflux}
\end{figure}

@@

\begin{table}[b!]
	\centering
	\caption{Peak neutron flux values from Moltres (this paper), COMSOL
	\cite{fiorina_molten_2013}, and OpenFOAM \cite{aufiero_development_2014}
	models along with the temperature distribution with which the values were
	obtained.}
	\begin{tabular}{l l S}
		\toprule
		{Model} & {Temperature distribution} & {Peak Neutron Flux [$\times 10^{15}$ cm$^{-2}$ s$^{-1}$]}
		\\
		\midrule
		{Moltres (This paper)} & {Steady state} & 9.80\\
		{COMSOL} & {Uniform, 973 K} & 8.6 \\
		{OpenFOAM} & {Uniform, 973 K} & 9.0 \\
		\bottomrule
	\end{tabular}
	\label{table:peak-flux}
\end{table}

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre1}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre2}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre3}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre4}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre5}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre6}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre7}
    \end{subfigure}
    \begin{subfigure}[t]{.243\textwidth}
        \centering
        \includegraphics[width=\textwidth]{ss-pre8}
    \end{subfigure}
    \caption{\gls{DNP} distributions in the core for \gls{DNP} groups
    1 to 8 (from left to right, top to bottom). Refer to Figure \ref{fig:neutronflux} for the height and radius
    scales on the y and x axes, respectively. Note the different scales for
    each distribution.}
    \label{fig:dnp}
\end{figure}

\subsection{Delayed Neutron Fraction}

As mentioned earlier, the delayed neutron precursors (DNPs) are mobile in
\glspl{MSR} and their distributions do not directly correspond to the neutron
flux distributions. The location where the \glspl{DNP} decay and emit neutrons
impacts their neutron
importance depending on their proximity to fissile and parasitic isotopes.
Figure \ref{fig:dnp} shows the \gls{DNP} distributions for all eight \gls{DNP}
groups. In general, we observe less \glspl{DNP} in the regions with fast salt
flow, namely
 The
precursors from the shortest-lived group (Group 8) predominantly decay within
the core as their half-lives are shorter than the time it takes to reach the
outlet while the precursors from the longest-lived group (Group 1) are
relatively evenly distributed due to their long half-lives. For the
longer-lived groups, the \gls{DNP} concentrations are not well-resolved on the
mesh elements adjacent to the outlet and the inlet boundaries, respectively.
Thus, we recommend careful mesh refinement for future work involving similar
geometries.

\begin{figure}[b!]
    \centering
    \begin{subfigure}[t]{.30\textwidth}
        \centering
        \vspace{.9cm}
        \includegraphics[width=\textwidth]{ss-pre}
    \end{subfigure}
    \begin{subfigure}[t]{.69\textwidth}
        \centering
        \vspace{0pt}
        \includegraphics[width=\textwidth]{fiorina-pre}
    \end{subfigure}
    \caption{Total \gls{DNP} distribution in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models.}
    \label{fig:pre}
\end{figure}

In Figure \ref{fig:pre}, we compare the total delayed neutron source
distribution from Moltres with the results from the Polimi and TUDelft models
\cite{fiorina_modelling_2014}. In contrast to Figure \ref{fig:dnp} which shows
the precursor distribution, Figure \ref{fig:pre} shows the rate of delayed
neutron emission, which we calculated by multiplying each \gls{DNP} group
$C_i$ with its associated decay constant $\lambda_i$.
We observe that the Polimi and TUDelft models
feature greater \gls{DNP} retention in the stagnant regions within the core.
This effect is less pronounced in the Moltres model. There is less build-up
of \glspl{DNP} at the top of the core in Moltres because most of the
\glspl{DNP} produced near the center of the core cannot enter the axial
recirculation zone that only appears in Moltres. We also observe less build-up
in the large recirculation zone near the blanket tank.

The in-core delayed neutron fraction $\beta_c$ is an important safety
parameter for \glspl{MSR}. This value represents the actual delayed
neutron fraction in \glspl{MSR} after accounting for the loss of delayed
neutrons from \glspl{DNP} decaying outside the active core region. In general,
reactors with smaller $\beta$ values exhibit greater prompt jumps in the
neutron flux in response to reactivity insertions because they have a greater
proportion of prompt neutrons under normal operating conditions. This is
undesirable from a reactor safety perspective because it exposes the reactor
to relatively more extreme conditions before active safety mechanisms
activate and scram the reactor. In \glspl{MSR}, this danger is partly
mitigated by the strong, negative fuel temperature reactivity coefficient. 
Chapter 6 contains more in-depth discussions for various transient scenarios.

In Table \ref{table:dnf}, we compare the fraction of out-of-core emissions and
the $\beta_c$ values from Moltres with the Polimi and TUDelft models. We
calculated the fraction of out-of-core emissions by finding the total amount
of each \gls{DNP} group in the core and the outer loop, multiplying each total
by their associated decay time constants $\lambda_i$, and calculating the
proportion of emissions in the outer loop relative to the grand total. We
calculated $\beta_c$ by first obtaining the prompt neutron emission rate from
Moltres and subsequently using in-core delayed neutron emission rate from the
previous calculation to find the fraction of delayed neutron emission rate
relative to total emission rate.

\begin{table}[t!]
	\centering
	\caption{The fraction of delayed neutrons lost from out-of-core emission
	and the in-core delayed neutron fraction $\beta_c$ values from Moltres
	(this paper), and the Polimi and TUDelft models
	\cite{fiorina_modelling_2014}.}
	\begin{tabular}{l S S}
		\toprule
		{Model} & {Out-of-Core Emission [\%]} & {$\beta_c$ [pcm]}
		\\
		\midrule
		{Moltres (This paper)} & {44.16} & {184.9}\\
		{Polimi} & {34.80} & {134.3} \\
		{TUDelft} & {34.85} & {123.8} \\
		\bottomrule
	\end{tabular}
	\label{table:dnf}
\end{table}

The fraction of out-of-core emissions from our Moltres model
differs significantly by approximately 10\%, while $\beta_c$ differs by
60-70 pcm. We attribute the former to the lesser \gls{DNP} retention in the
stagnant flow regions in the core; the \glspl{DNP} are more evenly distributed
along the entire primary loop, leading to more delayed neutron emissions in
the outer loop region. The most likely reason for this is differences in the
flow pattern in the recirculation zone because convective species transport
dominates diffusive effects in the \gls{MSFR}. The exact flow pattern in the
recirculation zones in the Polimi and TUDelft models is likely to differ from
that in Moltres (Figure \ref{fig:flow}). Figure \ref{fig:flow-temp} also shows
some minor differences in the magnitude of the flow in the recirculation zones
between Moltres, and the Polimi and TUDelft models. Although the sizes of the
arrows representing flow velocity are not normalized to the same scales, a
quick comparison between the largest arrows and the arrows in the
recirculation zone indicate that the recirculation zone in Moltres is
relatively more stagnant. This could explain the concentration of \glspl{DNP}
along an ``arc'' closer to the center of the core in Moltres as opposed to the
more even distribution of \glspl{DNP} throughout the whole recirculation zones
in the Polimi and TUDelft models. The higher peak \glspl{DNP} distribution in
Moltres also supports this assertion.

In spite of the greater delayed neutron losses, the $\beta_c$ value is higher
in Moltres than the Polimi and TUDelft models. To account for this
peculiarity, we note that Fiorina et al. \cite{fiorina_modelling_2014}
applied adjoint flux weighting for their $\beta_c$ calculation while we report
our value as the physical fraction without adjoint weighting. The weighting
results in a significant difference in $\beta_c$ because a large fraction of
the \glspl{DNP} decay in the recirculation zones, where the neutron importance
is noticeably diminished.
